Bip32-Ed25519++: A More Secure Hierarchical Deterministic Wallet Standard for Ed25519 Keys

### 2024-05-03 paper.tex
conclusion

\subsection{Hierarchical Deterministic Wallets}

@@

While reviewing the BIP32-Ed25519 paper, two issues were identified, the first minor and the other major.

@@

Truncate the 32 byte $Z$ to 28 bytes: $Z_L \leftarrow Z_{[:28]}$.

@@

Consider the scenario where a single child private key is compromised. The security of its parent key would degrade to only $\sqrt{2^{224}}=2^{112}$ due to the Birthday Attack.

If even a single child private key is compromised, then the security of its parent key degrades to only 112--114 bits. In other words, computing the parent key may become more than 64,000 times ($\sqrt{2^{32}}=2^{16}$) easier for the attacker as a result of doing away with 4 bytes of entropy.

@@

While the BIP32 approach cannot produce completely unrelated Ed25519 keys, our modifications ensure a reduction of only about 3 bits [DEPENDS ON WHICH "g" VALUE YOU CHOSE; IT'S (g-4)/2 BITS] of security in the above scenario.

This is done without sacrificing the number of child keys a parent can have, and by retaining a deep enough hierarchy of keys for applications.

@@

>>>>

\subsection{Hierarchical Deterministic Wallets}
\label{subsec:hd_wallets}

@@

While reviewing the BIP32-Ed25519 paper\cite{BIP32-Ed25519}, two issues were identified, the first minor and the other major.

@@

Truncate the 32 byte $Z$ to 28 bytes: $Z_L \leftarrow Z_{[:224]}$.

@@

Consider the scenario where a single child private key is compromised. The security of its parent key would degrade to around only $\sqrt{2^{224}}=2^{112}$ due to the Birthday Attack, below the 128 bits of security offered by Ed25519. This contradicts the author's claims that no loss of security occurs as a result of their method.

In other words, computing the parent key may become more than 64,000 times ($\sqrt{2^{32}}=2^{16}$) easier for the attacker as a result of doing away with 4 bytes of entropy.

@@

Define $g$, such that $Z_L = Z_{[:256-g]}$. The BIP32-Ed25519 paper\cite{BIP32-Ed25519} set $g=32$ to arrive on 28 bytes. An Ed25519 scalar has 251 bits available for randomization (255, 254, 2, 1, 0 are fixed), with an additional bit (253) having to be 0 to avoid overflow. The level at which we can no longer guarantee 253 will be 0 forms an upper bound.

We thus require, for security and correctness, that the following inequality holds:

\begin{equation*}
  g \geq d + 6
\end{equation*}

Where $2^d = D$, the number of levels of derivation. For best results, set the left hand and right hand sides equal.

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|c|}
  \hline
  g & d  & D  \\
  \hline
  8 & 2 & 4 \\
  9 & 3 & 8 \\
  16 & 10 & 1024 \\
  32 & 20 & 1048576 \\
  \hline
  \end{tabular}
  \caption{Bits to randomize $g$ resulting in levels $D$.}
  \label{tab:g_and_d}
  \end{table}

Table \ref{tab:g_and_d} presents a selection of possible values. $g=8$ is the most secure and breaks at the byte boundary (= 31 bytes) but results in only 4 levels of derivation - too few to satisfy the derivation schemes mentioned in Section \ref{subsec:hd_wallets} which require 5.

$g=16$ is the next value that breaks at the byte boundary (= 30 bytes) and provides more than enough levels. It randomizes 240 out of the 251 bits available in a Ed25519 scalar.

$g=9$ is the most secure value that also allows for enough derivation levels, allowing for 8 levels of derivation. It randomizes 246 out of the 251 bits available in a Ed25519 scalar. The only downside is that it does not break at the byte boundary, requiring a little bit of extra care when implementing.

@@

In particular, setting $g=9$ allows for the most secure implementation that also satisfies the requirements of the most common derivation schemes.

We have also provided implementations of this modification in several programming languages. We hope that this modification will be adopted by the wider community and help to improve the security of Ed25519-based hierarchical deterministic wallets.
